Mixed effect tweedie model in TMB

### 2019-07-13 analysis/tmb-sensor-explore/3-fit-do.Rmd
more plots and model configs for DO

  g <- plot_map(predictions, "est_rf") +
    ggtitle("Spatial random effects only") +
    scale_colour_gradient2()
  print(g)

@@

  g <- ggplot(predictions, aes(predictions[[variable]], residuals)) +
    geom_point() +
    geom_smooth() +
    facet_wrap(~year)
  print(g)

  g <- ggplot(predictions, aes(X, Y, colour = residuals)) +
    geom_point() +
    coord_fixed() +
    scale_color_gradient2()
  print(g)

@@

@@

```{r}
    starttime1 <- Sys.time()

    temp_model_all <- sdmTMB::sdmTMB(d_trawl1,
      temp_scaled ~ 0 + as.factor(year),
      time_varying = ~ 0 + depth_scaled + depth_scaled2,
      time = "year", spde = spde,
      family = gaussian(link = "identity"),
      ar1_fields = TRUE,
      include_spatial = TRUE,
      enable_priors = FALSE,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime1 <- Sys.time()
    time1 <- round(starttime1 - endtime1)

#saveRDS(do_model, file = (here::here("analysis/tmb-sensor-explore/models/model-do-1n3.rds")))
time1
temp_model_all
```

```{r}
get_diag(temp_model_all, response = "temp_scaled", variable = "depth_scaled")
```

Add temperature predictions to newdata grid for DO predictions
```{r}

temperature <- readRDS(here::here("analysis/VOCC/data/predicted_temp_allyears.rds"))
years <- unique(d_trawl1$year) 
temperature <- filter(temperature, year %in% years)
unique(temperature$year)

temperature$temp <- temperature$est
#temperature$temp_scaled2 <- temperature$est^2
temperature$raw_depth <- temperature$depth
temperature$depth <- log(temperature$depth)

@@

glimpse(newdata)

@@

```{r}
    starttime1 <- Sys.time()

    do_model_depth2 <- sdmTMB::sdmTMB(d_trawl1,
      log_do ~ 0 + as.factor(year) + temp_scaled + temp_scaled2,
      time_varying = ~ 0 + depth_scaled + depth_scaled2, #+ depth_scaled3,
      time = "year", spde = spde,
      family = gaussian(link = "identity"),
      ar1_fields = TRUE,
      include_spatial = TRUE,
      enable_priors = FALSE,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime1 <- Sys.time()
    time1 <- round(starttime1 - endtime1)

saveRDS(do_model_depth2, file = (here::here("analysis/tmb-sensor-explore/models/model-do-depth2.rds")))
time1
do_model_depth2
```

```{r}
do_model_DOY <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-all-DOY.rds"))

get_diag(do_model_DOY,response = "log_do", variable = "log_depth_scaled")

get_diag(do_model_DOY,response = "log_do", variable = "temp_scaled")
```

```{r}
do_model_depth3 <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-depth3.rds"))

get_diag(do_model_depth3,response = "log_do")

p_nd <- predict(do_model_depth3, newdata = newdata)

glimpse(p_nd)
```

```{r}
plot_map_raster <- function(dat, column = "est") {
  ggplot(dat, aes_string("X", "Y", fill = column)) +
    geom_raster() +
    facet_wrap(~year) +
    coord_fixed()
}

p_nd$do_est <- exp(p_nd$est)
hist(p_nd$do_est)
max(p_nd$do_est)

plot_map_raster(p_nd, "do_est") +
  scale_fill_viridis_c(trans = "log",) +
#  scale_fill_viridis_c(trans = "sqrt", option = "C") +
  ggtitle("Prediction (fixed effects + all random effects)")
```

```{r}
do_model1 <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-1.rds"))
get_diag(do_model1)
```

```{r}
plot_map_raster(p_nd, "est") +
  scale_fill_viridis_c(trans = "sqrt", option = "C") +
  ggtitle("Prediction (fixed effects + all random effects)")
```

```{r}
plot_map_raster(p_nd, "est_non_rf") +
  ggtitle("Prediction (fixed effects only)") +
  scale_fill_viridis_c(trans = "sqrt")
```

```{r, eval=FALSE}
#glimpse(p_nd)
plot_map_raster(p_nd, "omega_s") +
  ggtitle("Spatial random effects only") +
  facet_wrap(~year) +
  scale_fill_gradient2(low = "navy", mid = "white", high = "darkred")
```

```{r}
plot_map_raster(p_nd, "epsilon_st") +
  ggtitle("Spatiotemporal random effects only") +
  facet_wrap(~year) +
  scale_fill_gradient2(low = "navy", mid = "white", high = "darkred")
```

>>>>

  g <- plot_map(predictions, "est_rf") +
    ggtitle("All random effects only") +
    scale_colour_gradient2()
  print(g)

  g <- plot_map(predictions, "omega_s") +
    ggtitle("Spatial random effects only") +
    scale_colour_gradient2()
 print(g)

   g <- plot_map(predictions, "epsilon_st") +
    ggtitle("Spatiotemporal random effects only") +
    facet_wrap(~year) +
    scale_colour_gradient2()
 print(g)

@@

  g <- ggplot(predictions, aes_string(variable, "residuals")) +
    geom_point() +
    geom_smooth() +
    facet_wrap(~year)
  print(g)

  g <- ggplot(predictions, aes(X, Y, colour = residuals)) +
    geom_point() +
    coord_fixed() +
    scale_color_gradient2()+
    facet_wrap(~year)
  print(g)

@@

```{r}
ggplot(d_trawl, aes(temp, log_do, colour = DOY, size = sqrt(temp_range))) +
  geom_point() +
  facet_wrap(~year) +
  scale_color_viridis_c()
```

@@

Test temp model on this subset of data
DOY doesn't improve model
```{r eval=FALSE}
    starttime1 <- Sys.time()

    temp_model_DOY <- sdmTMB::sdmTMB(d_trawl1,
      temp_scaled ~ 0 + as.factor(year), # + DOY_scaled,
      time_varying = ~ 0 + depth_scaled + depth_scaled2,
      time = "year", spde = spde,
      family = gaussian(link = "identity"),
      ar1_fields = TRUE,
      include_spatial = TRUE,
      enable_priors = FALSE,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime1 <- Sys.time()
    time1 <- round(starttime1 - endtime1)

saveRDS(temp_model_all, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-2008-on.rds")))

# saveRDS(temp_model_DOY, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-2008-on-DOY.rds")))
# time1
# temp_model_DOY
temp_model_all
```

```{r eval=FALSE}
get_diag(temp_model_all, response = "temp_scaled", variable = "depth_scaled")
```

Add temperature predictions to newdata grid for DO predictions
```{r}
temperature <- readRDS(here::here("analysis/VOCC/data/predicted_temp_allyears.rds"))
years <- unique(d_trawl1$year) 
temperature <- filter(temperature, year %in% years)
unique(temperature$year)

temperature$temp <- temperature$est
#temperature$temp_scaled2 <- temperature$est^2
temperature$raw_depth <- temperature$depth
temperature$depth <- log(temperature$depth)
temperature$DOY_scaled <- 0

@@

glimpse(newdata)
```

Current best model of DO
```{r eval=FALSE}
    starttime1 <- Sys.time()

    do_model_re_DOY <- sdmTMB::sdmTMB(d_trawl1,
      log_do ~ 0 + as.factor(year) + temp_scaled + temp_scaled2 ,
      time_varying = ~ 0 + DOY_scaled + depth_scaled + depth_scaled2 + depth_scaled3,
      time = "year", spde = spde,
      family = gaussian(link = "identity"),
      ar1_fields = TRUE,
      include_spatial = TRUE,
      enable_priors = FALSE,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime1 <- Sys.time()
    time1 <- round(starttime1 - endtime1)

saveRDS(do_model_re_DOY, file = (here::here("analysis/tmb-sensor-explore/models/model-do-re-DOY.rds")))
time1
do_model_re_DOY
```

```{r}
do_model_re_DOY <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-re-DOY.rds"))
do_model_re_DOY
```

```{r}
get_diag(do_model_all_fe, response = "log_do", variable = "depth_scaled")

@@

```{r}
do_model_DOY <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-all-DOY.rds"))
do_model_DOY

get_diag(do_model_DOY, response = "log_do", variable = "depth_scaled")

# get_diag(do_model_DOY, response = "log_do", variable = "temp_scaled")
```

Other models tried:

```{r eval=FALSE}
do_model_DOY <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-DOY.rds"))
do_model_DOY
#p_nd <- predict(do_model_DOY, newdata = newdata)
get_diag(do_model_DOY, response = "log_do", variable = "depth_scaled")

```{r eval=FALSE}
do_model_depth2 <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-depth2.rds"))
do_model_depth2
get_diag(do_model_depth2,response = "log_do", variable = "depth_scaled")
#p_nd <- predict(do_model_depth2, newdata = newdata)
```

```{r eval = FALSE}
do_model_depth3 <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-depth3.rds"))
get_diag(do_model_depth3,response = "log_do")
#p_nd <- predict(do_model_depth3, newdata = newdata)
```

```{r eval=FALSE}
do_model_all <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-all.rds"))
#p_nd <- predict(do_model_all, newdata = newdata)
do_model_all
get_diag(do_model_all, response = "log_do", variable = "depth_scaled")
```

Predict DO for whole grid using quadratic of predicted temperature values and mean DOY
```{r}
p_nd <- predict(do_model_DOY, newdata = newdata)
p_nd$do_est <- exp(p_nd$est)

p_nd <- p_nd %>% filter(depth>15)
hist(p_nd$do_est)

max(p_nd$do_est)
max(d_trawl1$do_mlpl)
glimpse(p_nd)
min(p_nd$raw_depth)
min(d_trawl1$raw_depth)
p_nd[p_nd$do_est>8, ]

p_nd <- p_nd %>% filter(depth>15)
p_nd_filtered[p_nd_filtered$do_est>8, ]

saveRDS(p_nd, file = (here::here("analysis/tmb-sensor-explore/models/predicted-DO.rds")))
```

```{r}
plot_map_raster <- function(dat, column = "est") {
  ggplot(dat, aes_string("X", "Y", fill = column)) +
    geom_raster() +
    coord_fixed()
}

plot_map_raster(p_nd, "do_est") +
  scale_fill_viridis_c(trans = "log") +
  facet_wrap(~year) +
  ggtitle("Prediction (fixed effects + all random effects)")

```{r}
plot_map_raster(p_nd, "est_non_rf") +
  ggtitle("Prediction (fixed effects only)") +
  facet_wrap(~year) +
  scale_fill_viridis_c()
```

```{r}
plot_map_raster(p_nd, "est_rf") +
  ggtitle("Prediction (all random effects only)") +
  facet_wrap(~year) +
  scale_fill_viridis_c()
```

```{r, eval=FALSE}
#glimpse(p_nd)
plot_map_raster(p_nd, "omega_s") +
  ggtitle("Spatial random effects only") +
   scale_fill_gradient2( high = "Steel Blue 4",
  low = "Red 3")+
gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank())
```

```{r}
plot_map_raster(p_nd, "epsilon_st") +
  ggtitle("Spatiotemporal random effects only") +
  facet_wrap(~year) +
  scale_fill_gradient2( high = "Steel Blue 4",
  low = "Red 3")
```

```{r}
p_nd1 <- p_nd %>% filter(ssid!=16) %>% filter(ssid!=4)
plot_map_raster(p_nd1, "epsilon_st") +
  ggtitle("Spatiotemporal random effects only") +
  facet_wrap(~year) +
  scale_fill_gradient2( high = "Steel Blue 4",
  low = "Red 3", limits = c(-0.4,0.5)) +
gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank())
```
```{r}
p_nd4 <- p_nd %>% filter(ssid==4) #%>% filter(ssid!=4)
plot_map_raster(p_nd4, "epsilon_st") +
  ggtitle("Spatiotemporal random effects only") +
  facet_wrap(~year) +
  scale_fill_gradient2( high = "Steel Blue 4",
  low = "Red 3", limits = c(-0.4,0.5)) +
gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank())
```

```{r}
p_nd16 <- p_nd %>% filter(ssid==16) #%>% filter(ssid!=4)
plot_map_raster(p_nd16, "epsilon_st") +
  ggtitle("Spatiotemporal random effects only") +
  facet_wrap(~year) +
  scale_fill_gradient2( high = "Steel Blue 4",
  low = "Red 3", limits = c(-0.4,0.5)) +
gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank())
```
